---
title: "Anne Arundel County Population Estimate Forecast"
author: "Rick Fisher"
date: "September 4, 2018"
description: "This code was devloped to forecast future population estimates for Anne Arundel County out twenty years based on historical data" 
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load needed library files

```{r}
library(forecast)
library(ggplot2)
library(readxl)
```

# Import population data from Excel 

```{r}
aaPopR <- read_excel("C:/RStats/Forecast/Data/aaPop.xlsx")
```

# Preview first few columns of data to verify it was loaded into memory properly

```{r}
head(aaPopR)
```

# Plot the data

```{r}
aaPopPlot <- ggplot(aaPopR, aes(x = YEAR, y = POP)) +
  geom_point() +
  ggtitle("County Population Growth  1950-2017") +
  scale_x_continuous(name = "Year") +
  scale_y_continuous(name = "Population") +
  theme_bw()
print (aaPopPlot)
```

# Run a simple regression model plotting population vs year

```{r}
 aaPopReg <- lm(POP ~ YEAR, aaPopR)
```

# View regression model summary  

```{r}
 summary(aaPopReg)
```

# Visualize the regression model

```{r}
aaPopPlot <- ggplot(aaPopR, aes(x = YEAR, y = POP)) +
    geom_point(shape = 1) +
    geom_smooth(method = 'lm', se = FALSE) + 
    ggtitle("County Population Growth 1950-2017") +
    scale_x_continuous(name = "Year") +
    scale_y_continuous(name = "Population") +
    annotate("text", x=1985, y=300000, label = "R^2 == 0.983", parse=T) +
    theme_bw()
  print(aaPopPlot)
```

# Create a time series object 

```{r}
aaPopTs <- ts(aaPopR[2], start = c(1950,1), frequency = 1)
```

# Forecast a 20 year trend using Holt's method

```{r}
fcPopHolt <- holt(aaPopTs, h = 20)
autoplot(fcPopHolt)
checkresiduals(fcPopHolt)
summary(fcPopHolt)
```

# Validate the use of Holt's forecast method by comparing results to ETS.

```{r}
fitaaPop <- ets(aaPopTs)
summary(fitaaPop)
autoplot(forecast(fitaaPop))
checkresiduals(fitaaPop)
```
